---
title: "RWorkheet_nandin#4a.Rmd"
author: "Mary coolen nandin"
date: "2024-10-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
table <- data.frame(
Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5,
10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5,7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0,
74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0,64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F",
"M", "F", "M", "M", "M", "M", "F", "F", "M", "F","F", "M", "M", "F", "M", "M", "M", "M")
)
table
```
a. Describe the data.
The data frame Table contains three columns: Shoe_size, Height, and Gender.
Shoe_size: A numeric vector representing the shoe sizes of individuals. The values range from 5.0 to 13.0.
Height: A numeric vector representing the height of individuals in inches. The values range from 59.0 to 77.0
1
inches. Gender: A categorical variable indicating the gender of each individual, with values “M” for male and
“F” for female.

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
males <- subset(table, Gender == "M", select = c(Shoe_size, Height))
females <- subset(table, Gender == "F", select = c(Shoe_size, Height))
males
```
```{r}
females
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
MeanShoeSize <- mean(table$Shoe_size)
MeanShoeSize
```
```{r}
meanHeight <- mean (table$Height)
meanHeight
```
d. Is there a relationship between shoe size and height? Why?

Yes, there is a relationship between the shoe size and height of the respondents for the reason that thethe shoesize aligns to their height.

Figure 1: Household Data

Factors

A nominal variable is a categorical variable without an implied order. This means
that it is impossible to say that ‘one is worth more than the other’. In contrast,
ordinal variables do have a natural ordering.

Example:
```{r}
Gender <- c("M","F","F","M")
factor_Gender <- factor(Gender)
factor_Gender
```


2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

Consider data consisting of the names of months:
"March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August",

"July","December","August","August","September","November","February",|
April")
```{r}
months_vector <- c("March","April","January","November","January","September","October","September","November","January","November","November","February","May","August","July","December","August","August","September","April")
factor_months_vector <- factor (months_vector)
factor_months_vector
```


3. Then check the summary() of the months_vector and factor_months_vector. | Inter-
pret the results of both vectors. Are they both equally useful in this case?
```{r}
summary(months_vector)
```
```{r}
summary(factor_months_vector)
```
4. Create a vector and factor for the table below.
```{r}
factor_data <-c("East","West","North")
levels <- c(1, 4, 3)
```
```{r}
levels
```
```{r}
new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)
```

5. Enter the data below in Excel with file name = import_march.csv
```{r}
library(readr)
import_march <- read_csv("/cloud/project/import_march.csv")
```
```{r}
import_march
```