---
title: "RWorksheet_nandin#4b.Rmd"
author: "Mary coolen nandin"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Using Loop Function**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value
```{r}
matrixx <- matrix(0, nrow= 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) { 
matrixx[i, j] <- abs(i-j)
  }
}
print(matrixx)
```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}
rows <- 5

for (i in 1:rows) {
  cat(rep("*", i), "\n")
}

```


3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
start <- 10
a <- 0
b <- 1
repeat {
newfibonacci <- a + b
if (newfibonacci > 500) {
  break
}
if (newfibonacci >= start) {
cat(newfibonacci, "\n")
}
a <- b
b <- newfibonacci
}

```

**Using Basic Graphics (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result

```{r}
library(readr)
data <- read_csv("/cloud/project/shoesize.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
df <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,
                13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0,
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M",
             "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
female_subset <- subset(df, Gender == "F")
male_subset <- subset(df, Gender == "M")

num_females <- nrow(female_subset)
num_males <- nrow(male_subset)

num_females
num_males


```


c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.

```{r}
gender_counts <- c(Females = 14, Males = 14 )    

barplot(
  gender_counts,
  main = "Number of Males and Females in Household Data",  
  xlab = "Gender",                                         
  ylab = "Count",                                          
  col = c("lightblue", "grey"),                   
  legend.text = c("Females", "Males"),                     
  beside = TRUE                                           
)

```

5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous
60 10 5 25

a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}

categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
expenses <- c(60, 10, 5, 25)


percent_labels <- paste0(round((expenses / sum(expenses)) * 100), "%")


colors <- c("lightblue", "orange", "green", "purple")


pie(expenses, labels = percent_labels, col = colors, main = "Dela Cruz Family Monthly Expenses")
legend("topright", legend = categories, fill = colors)

```


6. Use the iris dataset.
data(iris)
a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.
```{r}
data(iris)
str(iris)

```


b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
mean_values <- colMeans(iris[, 1:4])
mean_values
```


c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.
```{r}
data(iris)
species_counts <- table(iris$Species)
pie(species_counts,
main = "Species Distribution in Iris Dataset",
col = c("grey", "#D3B", "#C17"),
labels = paste(names(species_counts), "\n", species_counts))
legend("topright", legend = names(species_counts), fill = c("grey", "#D3B", "#C17"))

```


d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
setosa_subset <- iris[iris$Species == "setosa", ]
versicolor_subset <- iris[iris$Species == "versicolor", ]
virginica_subset <- iris[iris$Species == "virginica", ]
last_six_setosa <- tail(setosa_subset, 6)
last_six_versicolor <- tail(versicolor_subset, 6)
last_six_virginica <- tail(virginica_subset, 6)
last_six_setosa

```
```{r}
last_six_versicolor
```
```{r}
last_six_virginica
```

e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
Male? How about in Female? Write the R scripts and its output.
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
main = "Iris Dataset",
sub = "Sepal Width and Length",
xlab = "Sepal Length",
ylab = "Sepal Width",
pch = 19,
col = iris$Species)
legend("topright", legend = levels(iris$Species),
col = 1:3, pch = 19)

```
f. Interpret the result.
The number of each iris species in the dataset is displayed in the pie chart. Every slice
shows the size of each group that determines the most prevalent species. The subsets show
the final six records for every species, enabling us to examine their particular measurements, such as sepal and petal sizes, and recognize their differences. The scatterplot displays the connection between sepal and width and length. In the event that setosa clusters at lower measures, this species often has smaller blooms, but if virginica and versicolor overlap, it indicates that their sizes are comparable, which makes it more difficult to distinguish between them.

BASIC CLEANING AND TRANSFORMATION OF OBJECTS

7. Import the alexa-file.xlsx. Check on the variations. Notice that there are ex- tra whitespaces among
black variants (Black Dot, Black Plus, Black Show, Black
Spot). Also on the white variants (White Dot, White Plus, White Show, White Spot).
a. Rename the white and black variants by using gsub() function.
```{r}
library(readxl)
alexa_data <- read_excel("/cloud/project/alexa_file.xlsx")
unique(alexa_data$variation)
```
```{r}
alexa_data$variation <- gsub("Black Dot", "BlackDot", alexa_data$variation)
alexa_data$variation <- gsub("Black Plus", "BlackPlus", alexa_data$variation)
alexa_data$variation <- gsub("Black Show", "BlackShow", alexa_data$variation)
alexa_data$variation <- gsub("Black Spot", "BlackSpot", alexa_data$variation)
alexa_data$variation <- gsub("White Dot", "WhiteDot", alexa_data$variation)
alexa_data$variation <- gsub("White Plus", "WhitePlus", alexa_data$variation)
alexa_data$variation <- gsub("White Show", "WhiteShow", alexa_data$variation)
alexa_data$variation <- gsub("White Spot", "WhiteSpot", alexa_data$variation)
unique(alexa_data$variation)

```
b. Get the total number of each variations and save it into another object. Save the object as variations.RData. Write the R scripts. What is its result? Hint: Use the dplyr package. Make sure to install
it before loading the package.
```{r}
library(dplyr)
```
```{r}
variations <- alexa_data %>%
count(variation)
print(variations)
```
```{r}
save(variations, file = "variations.RData")
```
SAMPLE OUTPUT

c. From the variations.RData, create a barplot(). Complete the details of the chart which include the
title, color, labels of each bar.
```{r}
library(dplyr)
load("variations.RData")
variations$variation <- gsub(" +", " ", variations$variation)
variations$variation <- trimws(variations$variation)
bardata <- variations$n
barnames <- variations$variation
barplot(
bardata,
main = "Counts of Variations",
col = rainbow(length(bardata)),
names.arg = barnames,
xlab = "Variations",
ylab = "Counts",
las = 2,
border = "grey"
)
text(
x = seq_along(bardata),
y = bardata + max(bardata) * 0.05,
labels = bardata,
pos = 3,
cex = 1,
col = "grey"
)

```
d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by side. Complete the
details of the chart.
```{r}
library(ggplot2)
library(dplyr)
load("variations.RData")
variations$variation <- gsub(" +", " ", variations$variation)
variations$variation <- trimws(variations$variation)
bwvariations <- variations %>%
filter(grepl("Black|White", variation))
bardata <- as.matrix(bwvariations$n)
barnames <- bwvariations$variation
barplot(
bardata,
beside = TRUE,
main = "Counts of Black and White Variations",
col = c("lightblue", "pink", "yellow", "grey"),
names.arg = barnames,
xlab = "Variations",
ylab = "Counts",
las = 2,
border = "black"
)
text(x = seq_along(bardata), y = bardata, labels = bardata, pos = 3, cex = 0.8, col = "black")
```
hel




