---
title: "RWorksheet_nandin#4b.Rmd"
author: "Mary coolen nandin"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Using Loop Function**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value
```{r}
matrixx <- matrix(0, nrow= 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) { 
matrixx[i, j] <- abs(i-j)
  }
}
print(matrixx)
```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure
```{r}
rows <- 5

for (i in 1:rows) {
  cat(rep("*", i), "\n")
}

```


3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
```

**Using Basic Graphics (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result

```{r}
library(readr)
data <- read_csv("/cloud/project/shoesize.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
df <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5,
                13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0,
             77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M",
             "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")
)
female_subset <- subset(df, Gender == "F")
male_subset <- subset(df, Gender == "M")

num_females <- nrow(female_subset)
num_males <- nrow(male_subset)

num_females
num_males


```


c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.

```{r}
gender_counts <- c(Females = 14, Males = 14 )    

barplot(
  gender_counts,
  main = "Number of Males and Females in Household Data",  
  xlab = "Gender",                                         
  ylab = "Count",                                          
  col = c("lightblue", "grey"),                   
  legend.text = c("Females", "Males"),                     
  beside = TRUE                                           
)

```

5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous
60 10 5 25

a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}

categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
expenses <- c(60, 10, 5, 25)


percent_labels <- paste0(round((expenses / sum(expenses)) * 100), "%")


colors <- c("lightblue", "orange", "green", "purple")


pie(expenses, labels = percent_labels, col = colors, main = "Dela Cruz Family Monthly Expenses")
legend("topright", legend = categories, fill = colors)

```


6. Use the iris dataset.
data(iris)
a. Check for the structure of the dataset using the str() function. Describe what you
have seen in the output.

b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?

c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write
the R script and its result.

d. Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.

e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
Male? How about in Female? Write the R scripts and its output.

